### 2019-12-13 docs/thesis-project/article.tex
thesis project update

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:relatedwork}
\subsection{Google File System}
Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers. In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures. Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception \cite{gfs}.  This assumption is typical in modern distributed systems of any kind. GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management. One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks \cite{gfs}. In the GFS architecture cluster consists of a single master and multiple chunk-servers. The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them. Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations. Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files. A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs. For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master.

@@

\newpage\subsection{Ceph File System}
Ceph \cite{ceph} is an open-source, storage system that, like GFS, runs over commodity hardware. Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure. The base architectural component of Ceph is the RADOS object store, and all content a client reads and writes to the system are represented as objects, irrespective of the objects' original data type. RADOS is for all purposes a Ceph cluster, and its storage space originates from the disks of multiple machines. Ceph architecture is slightly more complicated than that of GFS, so we will summarize the components that are of interest to us. Inside a cluster, we find three types of machines, storage devices, abstracted with a layer called Object Software Daemon (OSD), Monitors, and Metadata Servers. A small count of Monitors maintain cluster membership and state using PAXOS \cite{paxos} to make consensual decisions. Metadata Servers only required sharing the distributed file system, are used for POSIX semantics, e.g., directory hierarchy and file metadata such as owner and last accessed timestamps. Monitors can detect if metadata servers are down and launch new ones as required. Finally, OSDs store clients' data, and there are anywhere between dozens to thousands of these in a Ceph cluster.

The innovation Ceph brings is that the OSDs within a cluster are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks. When one OSD suspects a peer as died or joined the cluster, it consults the Monitors to obtain an updated cluster map and possibly offloads some of his data to some other living OSD. The way an OSD knows based on the cluster map if he needs to send data to another OSD is the same process a client goes through to discover to which OSD he should write or read his data. In all cases, a client or OSD process runs the Ceph.CRUSH(placement groups, cluster state, rule set) to obtain the Target OSDs of the named object. The last two parameters are obtained directly from the monitors. The first parameter is the result of the operation: hash(object name) mod (size of object pool). CRUSH \cite{crush} is a light-weight pseudo-random placement algorithm, that provides deterministic object quantization without lookups to central directories. Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings.

Ceph is proof that cloud systems can leverage P2P behavior in order to dynamically adjust replication in a decentralized fashion while meeting high dependability standards. \cite{ceph, ceph_benchmarks}

@@

An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i, j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.

@@

\subsection{Unstructured Overlays}
\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}. Gnutella is the first-ever decentralized, unstructured P2P protocol. Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation. Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive. Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations. The results show that Gia is three to five orders of magnitude better than that of Gnutella. Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent. Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.

@@

\newpage\section{Membership Management Protocols}
\subsection{CYCLON} CYCLON\cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays. CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner. One fundamental issue of unstructured networks is to avoid partitioning. One naive solution would be for each peer to keep a view of all the peers in the network and create connections has needed, but that solution is not scalable.

>>>>

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:relatedwork}
Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines. It is common practice to replicate data at the file-level, block-level, or both, depending on the requirements of each system. For example, block-level replication may allow parallel execution of applications, such as is the case of Hadoop, Google, and Spark MapReduce algorithms. Here we present two DFS, which follow entirely different approaches when it comes to architecture and protocols. We focus only on the architectural implementation, their advantages or disadvantages. We do not consider parallel computation because our primary objective, in this thesis proposal, is the continuous availability of data in a system, to clients. Before we dive into them, it is essential to understand some desired characteristics of a DFS. In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable. To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}). Features that contribute to the above characteristics are summarized below.

\textbf{\textit{Load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system. It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures. Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely,  when some part of it fails. As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.

\subsection{Google File System}

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{images/gfs.png}
\caption{GFS Architecture}\label{fig:gfs_arch}
\end{figure}

Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers. In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures. Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception \cite{gfs}.  This assumption is typical in modern distributed systems of any kind. GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management. One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks \cite{gfs}. In the GFS architecture cluster consists of a single master and multiple chunk-servers. The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them. Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations. Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files. A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs. For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master. See Fig.\ref{fig:gfs_arch} for a simplified architecture diagram.

@@

\newpage\subsection{Ceph File System}
Ceph \cite{ceph} is an open-source, storage system that, like GFS, runs over commodity hardware. Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure. The base architectural component of Ceph is the \textit{RADOS} object store. All content clients read and write to the system is represented as an object, irrespective of the object's original data type. \textit{RADOS} is a single logical entity composed of many \textit{Ceph Storage clusters}. Ceph architecture, Fig. \ref{fig:cfs_arch} is slightly more sophisticated than that of GFS, so we will summarize the components that are of interest to us. Inside a \textit{Ceph Storage cluster}, we find four types of logical machines, storage devices, abstracted with a layer called \textit{Object Software Daemon} (OSD), \textit{Monitors}, \textit{Managers}, and \textit{Metadata Servers}. \textit{Metadata Servers} are only required when a client his sharing his DFS with other clients, these servers contain information used for POSIX semantics, e.g., file metadata such as owner and last accessed timestamps. A small amount of \textit{Monitors} maintain a master copy of the storage cluster map with its current state. Monitors require high consistency and use Paxos \cite{paxos} to decide on the state of the cluster. \textit{Managers} maintains detailed information about placement groups, process metadata and host metadata. Finally, OSD store data on behalf of clients. Additionally, OSD utilize their nodes' resources to perform data replication, erasure coding, load-balancing, recovery, monitoring and reporting functions. Each Storage Cluster as anywhere from dozens to thousands of these OSDs.

The innovation Ceph brings is that the OSD within the same \textit{placement group}, within a cluster, are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks. \textit{Placement groups} are fragments of a logical \textit{pool}, and in turn, \textit{pools} are logical, dynamic partitions that define among other things, how much and which type of replication is done to objects that belong to it. When one OSD in a placement group suspects another as died or joined the cluster, it consults a \textit{Monitor} to obtain an up to date cluster map. Based on the cluster map and the \textit{CRUSH}\cite{crush} algorithm, it is possible that this OSD will have to offload some of its content to other OSD. Clients also use \textit{CRUSH} to perform reads and writes to the system. To perform a write operation, for example, a client connects to a \textit{pool} in the storage cluster and runs \textit{CRUSH} with a cluster map retrieved from a Monitor. The result of the operation gives the client information on the placement group of the data he wishes to upload as well as the OSD he should contact to complete his request. When the target OSD receives the request, it runs \textit{CRUSH} to find the number of replicas that it should store. Then, it takes the object identifier, pool name, and the cluster map and uses \textit{CRUSH}, again, to discover the secondary OSDs to where he will replicate the write. The operation is completed when the primary OSD, receives acknowledgments from the remaining secondary OSD and acknowledges back to the client. \textit{CRUSH} is a light-weight pseudo-random placement algorithm that provides deterministic object quantization without lookups to central directories. Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings. Ceph is proof that distributed file systems can reliably be implemented using P2P behavior. Ceph OSDs perform a lot of self monitoring and management operations automatically and in a decentralized fashion, while being highly dependable. \cite{ceph, ceph_benchmarks}

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{images/cephfs.png}
\caption{CFS Architecture \cite{configure-ceph}}
\label{fig:cfs_arch}
\end{figure}

@@

An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i}, \textit{j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.

@@

\newpage\subsection{Unstructured Overlays}
\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}. Gnutella is the first-ever decentralized, unstructured P2P protocol. Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation. Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive. Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations. The results show that Gia is three to five orders of magnitude better than that of Gnutella. Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent. Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.

@@

\section{Membership Management Protocols}
\subsection{CYCLON} CYCLON\cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays. CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner. One fundamental issue of unstructured networks is to avoid partitioning. One naive solution would be for each peer to keep a view of all the peers in the network and create connections has needed, but that solution is not scalable.
